---
title: "Space-mission"
author: "Phineas Pham"
date: '2022-10-16'
output: html_document
---

```{r setup, include=FALSE, warning = F}
knitr::opts_chunk$set(echo = TRUE)
require(astsa) #Library for book
require(xts)
require(mosaic)
require(dplyr)
require(car)
require(Stat2Data)
require(dynlm)
library(nlme)
require(AER)
library(forecast)
require(mgcv)
library(tseries) # need for Augmented Dickey-Fuller test
require(lmtest) # need for Durbin Watson test
require(fBasics) # need for normality tests of residuals
require(leaps)
require(urca) # need for ERS test of stationarity
library(tidyverse)
library(lubridate)
library(readr)
library(padr)
```

**Input data**

```{r, warning = FALSE}
space_missions <- read_csv("space_missions.csv")
#View(space_missions)
space_missions_data_dictionary <- read_csv("space_missions_data_dictionary.csv")
#View(space_missions_data_dictionary)
```


**Data Wrangling**

We create a new dataframe with 2 columns: 1 for months and the other for the number of space missions in that month.

```{r}
#group data by month and sum sales
sp <- space_missions %>% 
    group_by(Date) %>%
    summarise(total_count=n(),
            .groups = 'drop') %>% 
    as.data.frame()

```

```{r}
sp <- sp %>% 
    group_by(month = lubridate::floor_date(Date, 'month')) %>%
    summarize(sum = sum(total_count)) 

new_data <- pad(sp) 
new_data[is.na(new_data)] <- 0
sp <- new_data

```

```{r}
spSum = sp$'sum'
#View(sp)
summary(sp)
times <- ts(spSum)
plot.ts(times, ylab = 'Number of space missions')
#tsplot(spSum, ylab = 'Number of space missions')

```


**Non Parametric Trend**

Smooth to with frequency = 1
```{r}
SOI = ts(spSum, freq=1)
tsplot(SOI, col=8, ylab = 'Number of space missions') # the time scale matters (not shown)
lines(ksmooth(time(SOI), SOI, "normal", bandwidth=12), lwd=2, col=4)

```

Smooth with frequency = 4

```{r}
SOI = ts(spSum, freq=4)
tsplot(SOI, col=8, ylab = 'Number of space missions') # the time scale matters (not shown)
lines(ksmooth(time(SOI), SOI, "normal", bandwidth=12), lwd=2, col=4)
```
Smooth with frequency = 12

```{r}
SOI = ts(spSum, freq=12)
tsplot(SOI, col=8, ylab = 'Number of space missions') # the time scale matters (not shown)
lines(ksmooth(time(SOI), SOI, "normal", bandwidth=12), lwd=2, col=4)
```


```{r}
#Lowess
trend(spSum, lowess = TRUE, main='Space missions')

```

**Fitting SARIMA Model**

Now we check the detrended and differenced time series to see which one is stationary

```{r}

par(mfrow=2:1) # plot transformed data
tsplot(detrend(spSum), main="detrended" )
tsplot(diff(spSum), main="differenced" )
```

Differenced time series looks stationary.


```{r}
tsplot(diff(sp$sum))

adf.test(diff(sp$sum))
pp.test(diff(sp$sum))
kpss.test(diff(sp$sum))

```
```{r}
summary(ur.ers(diff(sp$sum)))
```

After 4 tests, I am confident that the differenced time series is stationary.

```{r, warning=FALSE}
mod1 <- Arima(sp$sum, order = c(0,1,0))
plot(mod1$residuals)
acf(mod1$residuals)
pacf(mod1$residuals)


mod1
```

```{r, warning=FALSE}
mod1 <- Arima(sp$sum, order = c(0,1,1), seasonal = list(order=c(2,1,2),period=12))
plot(mod1$residuals)
acf(mod1$residuals)
pacf(mod1$residuals)

adf.test(mod1$residuals)
pp.test(mod1$residuals)
kpss.test(mod1$residuals)
summary(ur.ers(mod1$residuals))

mod1
```
```{r, warning=FALSE}
mod1 <- Arima(sp$sum, order = c(1,2,1), seasonal = list(order=c(2,1,2),period=12))
plot(mod1$residuals)
acf(mod1$residuals)
pacf(mod1$residuals)

adf.test(mod1$residuals)
pp.test(mod1$residuals)
kpss.test(mod1$residuals)
summary(ur.ers(mod1$residuals))

mod1
```
```{r, warning=FALSE}
mod1 <- Arima(sp$sum, order = c(2,2,1), seasonal = list(order=c(2,2,2),period=12))
plot(mod1$residuals)
acf(mod1$residuals)
pacf(mod1$residuals)

adf.test(mod1$residuals)
pp.test(mod1$residuals)
kpss.test(mod1$residuals)
summary(ur.ers(mod1$residuals))

mod1
```


```{r}
auto.arima(sp$sum)
```
```{r, warning=FALSE}
mod1 <- Arima(sp$sum, order = c(1,1,2))
plot(mod1$residuals)
acf(mod1$residuals)
pacf(mod1$residuals)

adf.test(mod1$residuals)
pp.test(mod1$residuals)
kpss.test(mod1$residuals)
summary(ur.ers(mod1$residuals))

mod1
```

```{r, warning=FALSE}
mod1 <- Arima(sp$sum, order = c(1,1,2), seasonal = list(order=c(0,1,2),period=12))
plot(mod1$residuals)
acf(mod1$residuals)
pacf(mod1$residuals)

adf.test(mod1$residuals)
pp.test(mod1$residuals)
kpss.test(mod1$residuals)
summary(ur.ers(mod1$residuals))

mod1
```
```{r, warning=FALSE}
mod1 <- Arima(sp$sum, order = c(1,1,2), seasonal = list(order=c(0,0,2),period=12))
plot(mod1$residuals)
acf(mod1$residuals)
pacf(mod1$residuals)

adf.test(mod1$residuals)
pp.test(mod1$residuals)
kpss.test(mod1$residuals)
summary(ur.ers(mod1$residuals))

mod1
```
```{r, warning=FALSE}
mod1 <- Arima(sp$sum, order = c(1,1,2), seasonal = list(order=c(2,0,0),period=12))
plot(mod1$residuals)
acf(mod1$residuals)
pacf(mod1$residuals)

adf.test(mod1$residuals)
pp.test(mod1$residuals)
kpss.test(mod1$residuals)
summary(ur.ers(mod1$residuals))

mod1
```
```{r, warning=FALSE}
mod1 <- Arima(sp$sum, order = c(1,1,2), seasonal = list(order=c(1,0,1),period=12))
tsplot(mod1$residuals)
coeftest(mod1)
acf(mod1$residuals)
pacf(mod1$residuals)


mod1
```
```{r}
coeftest(Arima(sp$sum, order = c(1,1,2), seasonal = list(order=c(1,0,1),period=12), include.drift = FALSE))
```
My best model:

```{r, warning=FALSE}
mod1 <- Arima(sp$sum, order = c(0,1,2), seasonal = list(order=c(1,0,1),period=12))
tsplot(mod1$residuals)
coeftest(mod1)
acf(mod1$residuals)
pacf(mod1$residuals)

mod1
```
```{r}
#stationary tests for residuals
adf.test(mod1$residuals)
pp.test(mod1$residuals)
kpss.test(mod1$residuals)
summary(ur.ers(mod1$residuals))
```


Model utility tests:

```{r}
checkresiduals(mod1)
tsdiag(mod1) # looks good
mod1 = sarima(spSum,p=0,d=1,q=2,P=1,D=0,Q=1,S=12)

densityplot(as.numeric(mod1$residuals)) # these look uniform

summary(mod1)
coeftest(mod1)
```
```{r}
pred = forecast(mod1,h=60)
pred 
```
Forecasting:

```{r}
forecastArea <- forecast(mod1$fitted, h = 60)
plot(forecastArea,lwd=2,col="purple", main="Forecasts from ARIMA(0,1,2)(1,0,1)[12]", xlab="Time", ylab="Number of space missions") 
legend("topleft", legend=c("Past", "Future"), col=c("Purple", "Blue"), lty=1:2, cex=0.8) 
   
```


**Function of time**

```{r,warning=F}
sp$M = as.factor(month(sp$month))
sp$Y = as.factor(year(sp$month))
```

```{r}
modY=lm(spSum~sp$month, data=sp)
summary(modY)

ResidLinear=ts(modY$residuals)
plot(ResidLinear,cex=1.5,cex.lab=1.5,cex.axis=1.5,lwd=2,col="darkblue",xlab="t",ylab="Residual",main="Residuals - Linear Fit")
abline(0,0,col="red")
points(sp$month,modY$residuals,pch=19,col="darkblue")
```



```{r}
t = ts(sp$month)
fit = lm(spSum~ t + I(t^2) + cos(t*(2*pi/12)) + sin(t*(2*pi/12)) , na.action=NULL)

summary(fit)
```




```{r}
t = ts(sp$month)
fit = lm(sp$sum ~ I(t^2) + as.factor(sp$M)  , na.action=NULL)

summary(fit)
vif(fit)

```

```{r}
t = ts(sp$month)
fit = lm(sp$sum ~ as.factor(sp$M)  , na.action=NULL)

summary(fit)


```










